Patching Particle-Particle Particle-Mesh Method

### 2023-04-13 docs/papers/pp/pp.tex
FIXUP draft: a question
FIXUP draft: add complexity analysis

\usepackage{CJK}
\usepackage{microtype}
%\usepackage{savesym}
%  \savesymbol{lambdabar}
\usepackage{newpxtext,eulerpx}
%  \restoresymbol{newpx}{lambdabar}
\usepackage[T1]{fontenc}
\usepackage[margin=3cm]{geometry}
%\usepackage{setspace}
\usepackage{amsmath}
%  \allowdisplaybreaks
\usepackage{bm}
%\usepackage{mathrsfs}
%\usepackage{soul}
%\usepackage{tensor}
\usepackage{graphicx}
%\usepackage{booktabs}
%\usepackage{placeins}
%\usepackage{subcaption}
%\usepackage[caption=false]{subfig}  % revtex hates caption
\usepackage{hyperref}
  \hypersetup{
    colorlinks=true,
    linkcolor=BrickRed,
    filecolor=BrickRed,
    citecolor=LimeGreen,
    urlcolor=Cerulean,
  }
\usepackage{natbib}
\usepackage[dvipsnames,svgnames,x11names]{xcolor}
\usepackage{comment}
  \newif\ifshow
  %\showtrue  % show
  \showfalse  % hide
  %
  \ifshow
    \includecomment{hide}
  \else
    \excludecomment{hide}
  \fi
\usepackage{tikz}
  \usetikzlibrary{positioning, fit, calc, arrows.meta}

@@

\newcommand{\order}{\mathcal{O}}
\newcommand{\deltaD}{\delta^\textsc{d}}
\newcommand{\deltaK}{\delta^\textsc{k}}
\renewcommand{\d}{d}
\newcommand{\p}{\partial}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cR}{\mathcal{R}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cH}{\mathcal{H}}
\bmdefine{\vzero}{0}
\bmdefine{\vI}{I}
\bmdefine{\vnabla}{\nabla}
\bmdefine{\vtheta}{\theta}  % parameters
\bmdefine{\vomega}{\omega}  % white noise modes
\bmdefine{\vk}{k}  % wavevectors
\bmdefine{\vn}{n}  % wavevector indices
\bmdefine{\vx}{x}  % comoving and canonical coordinates
\bmdefine{\vq}{q}  % Lagrangian coordinates
\bmdefine{\vs}{s}  % displacements
\bmdefine{\vp}{p}  % canonical momenta
\bmdefine{\va}{a}  % accelerations
\bmdefine{\vz}{z}  % states
\bmdefine{\vf}{f}
\bmdefine{\vF}{F}
\bmdefine{\vvarphi}{\varphi}
\newcommand{\As}{A_\mathrm{s}}
\newcommand{\ns}{n_\mathrm{s}}
\newcommand{\Omegam}{\Omega_\mathrm{m}}
\newcommand{\Omegab}{\Omega_\mathrm{b}}
\newcommand{\Mpc}{\mathrm{Mpc}}
\newcommand{\ic}{\mathrm{i}}
\newcommand{\knyq}{k_\mathrm{Nyq}}
\newcommand{\kfun}{k_\mathrm{fun}}

@@

\YL{TODO: verify that in 1D the partial sum explains the force error.}

@@

>>>>

\usepackage{CJK}
\usepackage{microtype}
%\usepackage{savesym}
%  \savesymbol{lambdabar}
\usepackage{newpxtext,eulerpx}
%  \restoresymbol{newpx}{lambdabar}
\usepackage[T1]{fontenc}
\usepackage[margin=3cm]{geometry}
%\usepackage{setspace}
\usepackage{amsmath}
%  \allowdisplaybreaks
\usepackage{bm}
\usepackage{dsfont}
%\usepackage{mathrsfs}
%\usepackage{soul}
%\usepackage{tensor}
\usepackage{graphicx}
%\usepackage{booktabs}
%\usepackage{placeins}
%\usepackage{subcaption}
%\usepackage[caption=false]{subfig}  % revtex hates caption
\usepackage{hyperref}
  \hypersetup{
    colorlinks=true,
    linkcolor=BrickRed,
    filecolor=BrickRed,
    citecolor=LimeGreen,
    urlcolor=Cerulean,
  }
\usepackage{natbib}
\usepackage[dvipsnames,svgnames,x11names]{xcolor}
\usepackage{comment}
  \newif\ifshow
  %\showtrue  % show
  \showfalse  % hide
  %
  \ifshow
    \includecomment{hide}
  \else
    \excludecomment{hide}
  \fi
\usepackage{tikz}
  \usetikzlibrary{positioning, fit, calc, arrows.meta}

@@

\DeclareMathOperator*{\E}{\mathds{E}}
\DeclareMathOperator{\tsum}{{\textstyle\sum}}
\newcommand{\1}{\mathds{1}}
\newcommand{\order}{\mathcal{O}}
\newcommand{\deltaD}{\delta^\textsc{d}}
\newcommand{\deltaK}{\delta^\textsc{k}}
\renewcommand{\d}{d}
\newcommand{\p}{\partial}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cR}{\mathcal{R}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cH}{\mathcal{H}}
\bmdefine{\vzero}{0}
\bmdefine{\vI}{I}
\bmdefine{\vnabla}{\nabla}
\bmdefine{\vtheta}{\theta}  % parameters
\bmdefine{\vomega}{\omega}  % white noise modes
\bmdefine{\vk}{k}  % wavevectors
\bmdefine{\vn}{n}  % wavevector indices
\bmdefine{\vx}{x}  % comoving and canonical coordinates
\bmdefine{\vq}{q}  % Lagrangian coordinates
\bmdefine{\vs}{s}  % displacements
\bmdefine{\vp}{p}  % canonical momenta
\bmdefine{\va}{a}  % accelerations
\bmdefine{\vz}{z}  % states
\bmdefine{\vf}{f}
\bmdefine{\vF}{F}
\bmdefine{\vvarphi}{\varphi}
\newcommand{\As}{A_\mathrm{s}}
\newcommand{\ns}{n_\mathrm{s}}
\newcommand{\Omegam}{\Omega_\mathrm{m}}
\newcommand{\Omegab}{\Omega_\mathrm{b}}
\newcommand{\Mpc}{\mathrm{Mpc}}
\newcommand{\ic}{\mathrm{i}}
\newcommand{\knyq}{k_\mathrm{Nyq}}
\newcommand{\kfun}{k_\mathrm{fun}}

@@

\YL{TODO: verify that in 1D the partial sum explains the force error.
Two ways: mass integral vs force partial sum, off by the Nyquist?}

@@

\subsection{Complexity}

$N$ particles, $M$ cells, volume $V$, mean matter density
$\bar\rho$, particle mass $m = \bar\rho V / N$, density field $\rho(\vx)
= \bar\rho [1 + \delta(\vx)]$, $C_i$ is the $i$-th cell, and $D_i$ are
the cells whose particles directly interact with those in $C_i$ (so $C_i
\subsetneq D_i$).
Each cell $C$ has volume $V / M$, and each direct interacting region $D$
has volume $d V / M$ with $d$ being an integer greater than 1.
The total number of pairs of interactions is
%
\begin{align}
\tsum_{i=1}^M \E \bigl[ \tsum_{p\in C_i} \1_{p'\in D_i} \bigr]
%
&\simeq \frac{N^2}{V^2} \tsum_{i=1}^M \E \biggl[
  \int_{C_i} \!\!\!\d\vx \int_{D_i} \!\!\!\d\vx'\,
  [1 + \delta(\vx)] [1 + \delta(\vx')] \biggr] \nonumber\\
%
&= \frac{M N^2}{V^2} \int_C \!\!\d\vx \int_D \!\!\d\vx'\,
  \bigl[ 1 + \xi(\vx' - \vx) \bigr]
= \frac{d N^2}M  \bigl( 1 + \sigma^2_{CD} \bigr),
\end{align}
%
where $\xi$ is the matter density 2-point correlation function, and we
have replaced the sums over particles with the integrals over the
density fields, $\sum_p \to m^{-1} \int \!\d\vx\, \rho(\vx) = N V^{-1}
\int \!\d\vx\, [1 + \delta(\vx)]$, in deriving the first equality.
We have also used the density covariance between $C$ and $D$, defined as
%
\begin{equation}
\sigma^2_{CD} \triangleq \int \!\!\frac{\d\vk}{(2\pi)^3}\,
  P(k) W_C^*(\vk) W_D(\vk)
= \frac{M^2}{d V^2} \int_C \!\!\d\vx \int_D \!\!\d\vx'\,
  \xi(\vx' - \vx),
\end{equation}
%
where $P(k)$ is the matter density power spectrum and the Fourier
transform of $\xi$.
$\sigma^2_{CD}$ depends on the geometry of $C$ and $D$, through their
Fourier space windows:
%
\begin{align}
W_C(\vk) &\triangleq \frac{M}V \int_C \!\!\d\vx\, e^{i\vk\cdot\vx},
  \nonumber\\
W_D(\vk) &\triangleq \frac{M}{dV} \int_D \!\!\d\vx\, e^{i\vk\cdot\vx},
\end{align}
%
both normalized to 1 at $\vk=\vzero$.
